Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update introduction

Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using the classical linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or even contains outliers, there is cause for concern. Conducting a goodness of fit test is an alternative method for inspecting the distribution, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

To scrutinize the relationship between residuals, predicted values and each of the explanatory variables, it is recommend to plot them on a scatter plot. In general, the model is well-behaved if there is any visual discoverable pattern. However, it is a very difficult task for a human judge, though to make a decision that there's nothing there. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. Generally, one looks for departures from nothingness like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and also use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

This is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the commonly used diagnostic plots, the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers. Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure [ref here], the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.

There are many types of departures from a good residual plot. Three of them will be considered in this paper, which are also the most commonly checked departures for a residual plot. The first one is non-linearity. Any unnatural and non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. At the top-right of Figure [ref here], there is an example residual plot of this type of departure. The second type of departure that will be considered is heterskedasticity. The usual pattern is the inconsistent distribution of the residuals for different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis. The last type of departure we will consider is non-normality. This departure is usually harder to be detected from the residuals plot compared to non-linearity and heterskedasticity. In most cases, a quantile-quantile plot can do a better job. Depending on the actual distribution of the residuals, different visual patterns will be discovered. Readers of the residuals plot will notice the data points do not distribute noramlly on the y-axis. For example, if the distribution is skewed, one will see the "tail" of the distribution as some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].

>>>>

Residuals might be plotted, as a histogram or quantile-quantile plot to examine the distribution. Using the classical linear regression model as an example, if the distribution is symmetric and unimodal, it is well-behaved. But if the distribution is skewed, bimodal, multimodal, or contains outliers, there is cause for concern. The distribution could also be inspected by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

Plotting the residuals against predicted values and each of the explanatory variables on a scatter plot is a recommend way to scrutinize their relationships. If there is any visually discoverable patterns, the model is potentially misspecified. However, it is a very difficult task for a human judge, though to make a decision that there's nothing there. It is especially common, particularly among new data analysts to report patterns when an experienced data analyst might quickly conclude that there are none. Generally, one looks for departures from nothingness like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

This is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the the types of departures that one expects to detect, and describes a formal process for reading residual plots, called visual inference, that can avoid the concerns about subjectiveness of human readers. Section \ref{experimental-design} describes the experimental setup to enable a comparison between decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure [ref here], the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. We will only discuss three most commonly checked departures in this paper. Namely, non-linearity, heterskedasticity and non-normality.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. At the top-right of Figure [ref here], there is a residual plot giving an example of the visual pattern of non-linearity when a cubic term is not captured by the regression model.

Heterskedasticity refers to the existence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.

Non-normality is usually harder to be detected from a residual plot compared to non-linearity and heterskedasticity. Besides, a quantile-quantile plot can often do a better job for this task. Note that not all regression models assume normality of the error term, but a certain amount of them does. If this assumption happens to be false, then it is expected to see data points do not distribute noramlly on the y-axis. For example, given a skewed error distribution, one will see fewer data points and more outliers on one side of the zeroline as shown in the bottom-right of Figure [ref here].

### 2022-09-06 paper_comparison/paper_comparison/paper_comparison.tex
update introduction

Residuals might be plotted, as a histogram or quantile-quantile plot to
examine the distribution. Using the classical linear regression model as
an example, if the distribution is symmetric and unimodal, it is
well-behaved. But if the distribution is skewed, bimodal, multimodal, or
even contains outliers, there is cause for concern. Conducting a
goodness of fit test is an alternative method for inspecting the
distribution, such as the Shapiro-Wilk Normality test
\citep{shapiro1965analysis}.

To scrutinize the relationship between residuals, predicted values and
each of the explanatory variables, it is recommend to plot them on a
scatter plot. In general, the model is well-behaved if there is any
visual discoverable pattern. However, it is a very difficult task for a
human judge, though to make a decision that there's nothing there. It is
especially common, particularly among new data analysts to report
patterns when an experienced data analyst might quickly conclude that
there are none. Generally, one looks for departures from nothingness
like non-linear dependency or heteroskedasticity. It is also possible to
conduct hypothesis tests for non-linear dependence
\citep{ramsey_tests_1969}, and also use a Breusch-Pagan test
\citep{breusch_simple_1979} for heteroskedasticity.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure {[}ref here{]}, the
top-left panel is a good residual plot with residuals evenly distributed
at both sides of the horizontal zero line showing no noticeable
patterns.

There are many departures from a good residual plot, but we will
consider only three types of them in this paper, which are also the most
commonly checked departures for a residual plot. The first one is
non-linearity. Any unnatural non-linear curve presented in the residual
plot could be considered an indicative of non-linearity. At the
top-right of Figure {[}ref here{]}, there is an example residual plot of
this type of departure. The second type of departure we will consider is
heterskedasticity. The usual pattern of this type of departure is the
inconsistent distribution of the residuals for different x values.
Visually, it usually results in the so-called ``butterfly'' shape as
shown in the bottom-left panel of Figure {[}ref here{]}, or the
``triangle'' shape where the smallest variance occurs at the edges of
the x-axis. The last type of departure is non-normality. This departure
is usually harder to be detected from the residuals plot compared to
non-linearity and heterskedasticity. In most the case, a
quantile-quantile plot can do a better job. Depending on the actual
distribution of the residuals, different visual patterns will be
discovered. If the distribution is skewed, one will see some outliers
presented on only one side of the zero line as shown in the bottom-right
of Figure {[}ref here{]}.

>>>>

Residuals might be plotted, as a histogram or quantile-quantile plot to
examine the distribution. Using the classical linear regression model as
an example, if the distribution is symmetric and unimodal, it is
well-behaved. But if the distribution is skewed, bimodal, multimodal, or
contains outliers, there is cause for concern. The distribution could
also be inspected by conducting a goodness of fit test, such as the
Shapiro-Wilk Normality test \citep{shapiro1965analysis}.

Plotting the residuals against predicted values and each of the
explanatory variables on a scatter plot is a recommend way to scrutinize
their relationships. If there is any visually discoverable patterns, the
model is potentially misspecified. However, it is a very difficult task
for a human judge, though to make a decision that there's nothing there.
It is especially common, particularly among new data analysts to report
patterns when an experienced data analyst might quickly conclude that
there are none. Generally, one looks for departures from nothingness
like non-linear dependency or heteroskedasticity. It is also possible to
conduct hypothesis tests for non-linear dependence
\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
\citep{breusch_simple_1979} for heteroskedasticity.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure {[}ref here{]}, the
top-left panel is a good residual plot with residuals evenly distributed
at both sides of the horizontal zero line showing no noticeable
patterns. There are various types of departures from a good residual
plot. We will only discuss three most commonly checked departures in
this paper. Namely, non-linearity, heterskedasticity and non-normality.

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values presented
in the residual plot could be considered as an indicative of
non-linearity. At the top-right of Figure {[}ref here{]}, there is a
residual plot giving an example of the visual pattern of non-linearity
when a cubic term is not captured by the regression model.

Heterskedasticity refers to the existence of nonconst error variance in
a regression model. It is mostly due to the strict but false assumptions
on the variance-covariance matrix of the error term. The usual pattern
of heterskedasticity on a residual plot is the inconsistent spread of
the residuals at different x values. Visually, it sometimes results in
the so-called ``butterfly'' shape as shown in the bottom-left panel of
Figure {[}ref here{]}, or the ``left-triangle'' and ``right-triganle''
shape where the smallest variance occurs at the edges of the x-axis.

Non-normality is usually harder to be detected from a residual plot
compared to non-linearity and heterskedasticity. Besides, a
quantile-quantile plot can often do a better job for this task. Note
that not all regression models assume normality of the error term, but a
certain amount of them does. If this assumption happens to be false,
then it is expected to see data points do not distribute noramlly on the
y-axis. For example, given a skewed error distribution, one will see
fewer data points and more outliers on one side of the zeroline as shown
in the bottom-right of Figure {[}ref here{]}.
